### 2023-02-20 Conference Submission/DraftDoc.tex
more edits 3

Nodes within the network influence eachother and so we might wish to draw on the rich literature of differential equation modelling, which have been useful in modelling inumerable dynamical systems. Attempting to directly model the temporal adjacency matrix presents another challenge; changes in the network are discrete jumps, not smooth curves which can be nicely modelled by differential equations.

We propose using random dot product graphs, which are well established in statistics, and singular value decomposition to resolve this issue[citation]. Creating a singular value decomposition of each of our network observations allows us to view the temporal network as a sample, from which we can model the dynamics of edge existence probabilities across time, now a continuous function. Because the progression of the system can now be viewed as a continuous transformation, we can now look to model the change of this system. To regenerate the network from the embedding, we can simply multiply our embedding matrices together. This will give us a random dot product graph, where the entries (i,j) of the matrix will be the probability of an edge from i to j existing[citation]. In this form, the problem can be viewed as a dynamical system We believe this approach will be useful when applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling.

By using this process, we also open the possibility of using the endlessly flexible toolbox of machine learning. In this paper we explore the usefulness of neural ODEs, a neural network that is trained to differential equation of temporal data, for discovering the governing function of a set of simple, synthetic temporal networks. This results in a black box DE, which is not very helpful for understanding our systems. To give more insight into what is really going on, we then use symbolic regression to find a function whose output is similar to that of the neural network. By doing this we may further understand the important factors that govern our temporal network.

>>>>

XXX Nodes within the network influence each other and XXX so we might wish to draw on the rich mathematical literature of differential equation modelling.

In the common representation of networks as binary-valued adjacency matrices, the events recorded in a temporal sequence of networks correspond to topological events, such as the appearence or the disappearance of link.
Because of the discrete nature of events, directly modelling the temporal networks as dynamical systems would require to handle discrete jumps.
The topological nature of temporal networks, and the discontinuous character of their temporal evolution, make it challenging to use differential equations techniques.

Here, we overcome the discretness problem by interpreting networks as Random Dot Product Graphs; a well established statistical model for complex networks, that embeds nodes in a low-dimensional metric space[citation]. In this way we translate the hard problem of modelling discrete events in the space of networks to the easier problem of modelling continous change in the embedding space. Then, we define and use systems of Neural Network Differential Equations (NNDE) to approximate the time evolution of the embedding space, and symbolic regression techniques to discover the functional form of the fitted NNDEs. These functional forms are interpretable (as they read as classic differential equations) and allow to predict forward in time the evolution of the temporal networks.

In this manuscript, we prove that the temporal network prediction problem can be successfully re-interpreted as a dynamical system modelling problem. In particular, we apply our proposed framework to XXX
The framework we are introducing is extremely flexible, and our research regarding the optimal structure of the Neural Networks used for the NNDEs is just started.
We are confident that future research can identify more fitting Neural Network structures than the simple one adopted here.
For this reason, we did not yet attempt to benchmark our model against other classic temporal network prediction methods.
As it is completely general, we believe that the framework we are introducing can be usefully applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling. In particular, we discuss how specific domain knowledge relative to the prediction scenario can be taken into account, moving from NNDEs to Universal Differential Equations.
